---
title: "Clustering"
author: "Faith"
date: "9/8/2020"
output:
  pdf_document: default
  html_document: default
---

Installing all the necessary packages
Creating a fuction to help load the data
```{r}
# cleam unnecessary items

gc()
rm(list = ls(all = TRUE))

packages<-function(x){
  x<-as.character(match.call()[[2]])
  if (!require(x,character.only=TRUE)){
    install.packages(pkgs=x,repos="http://cran.r-project.org")
    require(x,character.only=TRUE)
  }
}

packages(corrplot)
packages(gridExtra)
packages(GGally)
packages(cluster) # clustering algorithms 
packages(factoextra) # clustering algorithms & visualization
```
Load the data
```{r}
setwd('C:/Users/FGakori/Documents/supervised and unsupervised')
wines <- read.csv('wine.csv')
```

```{r}
head(wines)
```
Removing the first column
```{r}
```
# Data Analysis

```{r}
summary(wines)
```

```{r}
str(wines)
```
The variables are either numerical / integers

```{r}
library(magrittr)
```


plot histogram for each attribute
```{r}
 
```

Correlation matrix

```{r}
corrplot(cor(wines), type = 'upper', method = 'number', tl.cex = 0.9)
```
There is a strong correlation between total_phenols and flavanoids. We can model the relationship between these two variables by fitting a linear equation.

```{r}
#  relationship btwn phenols and flavanoids

ggplot(wines, aes(x = Phenols, y = Flavanoids)) +
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE) +
  theme_bw()
```

# preparing data for k-means

normalize the data

```{r}
winesNorm <- as.data.frame(scale(wines))
summary(winesNorm)
```
Computing k-means clustering in R

We can compute k-means in R with the kmeans function. Here will group the data into two clusters (centers = 2). The kmeans function also has an nstart option that attempts multiple initial configurations and reports on the best one. For example, adding nstart = 25 will generate 25 initial configurations and reports on the best one. For example, adding nstart=25 generates 25 initial configurations. This approach is often recommended.

```{r}
set.seed(123)
wines_k2 <- kmeans(winesNorm, centers = 2, nstart = 25)
print(wines_k2)
```
visualize cluster created

```{r}
fviz_cluster(wines_k2, data = winesNorm)
```
When we print the model we build (wines_k2), it shows information like, number of clusters, centers of the clusters, size of the clusters and sum of square. Let’s check how to get these attributes of our model.

```{r}
# cluster to whih each point is associated
wines_k2$cluster
```
```{r}
# cluster centers(means)

wines_k2$centers
```
```{r}
# cluster size
wines_k2$size
```
```{r}
# between clusters sum of squares
wines_k2$betweenss
```

```{r}
# total sum of squares
wines_k2$tot.withinss
```

```{r}
# total sum of squares
wines_k2$totss
```

the number of clusters (k) must be set before we start the algorithm, it is often advantageous to use several different values of k and examine the differences in the results.

We can execute the same process for 3, 4, and 5 clusters, and the results are shown in the figure:

```{r}
wines_K3 <- kmeans(winesNorm, centers = 3, nstart = 25)
wines_K4 <- kmeans(winesNorm, centers = 4, nstart = 25)
wines_K5 <- kmeans(winesNorm, centers = 5, nstart = 25)
```

plot the clusters to compare different k values
```{r}
p1 <- fviz_cluster(wines_k2, geom = "point", data = winesNorm) + ggtitle(" K = 2")
p2 <- fviz_cluster(wines_K3, geom = "point", data = winesNorm) + ggtitle(" K = 3")
p3 <- fviz_cluster(wines_K4, geom = "point", data = winesNorm) + ggtitle(" K = 4")
p4 <- fviz_cluster(wines_K5, geom = "point", data = winesNorm) + ggtitle(" K = 5")

grid.arrange(p1, p2, p3, p4, nrow = 2)
```
# Determine optimal clusters

K-means clustering requires that you specify in advance the number of clusters to extract. A plot of the total within-groups sums of squares against the number of clusters in a k-means solution can be helpful. A bend in the graph can suggest the appropriate number of clusters.

Below are the methods to determine the optimal number of clusters

Elbow method
Silhouette method
Gap statistic


```{r}
# method 1
# determine optimal clusters(k) using elbow method

fviz_nbclust(x = winesNorm,FUNcluster = kmeans, method = 'wss' )
```

creating a function 
```{r}
wssplot <- function(data, nc = 15, set.seed = 1234){
  wss <- (nrow(data) - 1)*sum(apply(data, 2, var))
  for(i in 2:nc) {
    set.seed(1234)
    wss[i] <- sum(kmeans(x = data, centers = i, nstart = 25)$withinss)
  }
  plot(1:nc, wss, type = 'b', xlab = 'Number of Clusters', ylab = 'Within Group Sum of Square',
       main = 'Elbow Method Plot to Find Optimal Number of Clusters', frame.plot = T,
       col = 'blue', lwd = 1.5)
}

wssplot(winesNorm)
```


# Determining Optimal clusters (k) Using Average Silhouette Method

```{r}
fviz_nbclust(x = winesNorm,FUNcluster = kmeans, method = 'silhouette' )
```
# There is another method called Gap-Static used for finding the optimal value of K.

```{r}
# compute gap statistic
set.seed(123)
gap_stat <- clusGap(x = winesNorm, FUN = kmeans, K.max = 15, nstart = 25, B = 50 )

# Print the result
print(gap_stat, method = "firstmax")
```



```{r}
# plot the result to determine the optimal number of clusters.
fviz_gap_stat(gap_stat)
```
Final analysis using three clusters

```{r}
# compute k-means clustering with k=3
set.seed(123)
final <- kmeans(winesNorm, centers = 3, nstart = 25)
print(final)
```

visualize the results
```{r}
fviz_cluster(final, data = winesNorm)
```

We can extract the clusters and add to our initial data to do some descriptive statistics at the cluster level

```{r}
winesNorm %>% 
  mutate_if(Cluster = final$cluster) %>%
  group_by(Cluster) %>%
  summarize_all('median')
```

# IRIS DATASET

```{r}
require('datasets')
```


```{r}
data('iris')

str(iris)
```


```{r}
summary(iris)
```


```{r}
head(iris)
```


```{r}
# remove species and store it in another variable
# normalize

iris.new <- iris[, c(1,2,3,4)]
iris.class <- iris[, 'Species']
head(iris.new)
```


```{r}
#previewing class columns
head(iris.class)
```


```{r}
# normalize
normalize <- function(x){
  return((x-min(x)) / (max(x) - min(x)))
}

iris.new$Sepal.Length<- normalize(iris.new$Sepal.Length)
iris.new$Sepal.Width<- normalize(iris.new$Sepal.Width)
iris.new$Petal.Length<- normalize(iris.new$Petal.Length)
iris.new$Petal.Width<- normalize(iris.new$Petal.Width)
head(iris.new)
```


```{r}
# apply kmeans
# centroids(k)=3

result <- kmeans(iris.new, 3)

# preciewing no. of records in each cluster
result$size
```


```{r}
# Getting the value of cluster center datapoint value(3 centers for k=3)

result$centers
```


```{r}
# Getting the cluster vector that shows the cluster where each record falls

result$cluster
```



```{r}
# Plotting to see how Sepal.Length and Sepal.Width data points have been distributed in clusters
plot(iris.new[c(1,2)], col = result$cluster)
```


```{r}
# Plotting to see how Sepal.Length and Sepal.Width data points have been distributed  
# originally as per "class" attribute in dataset

plot(iris.new[c(1,2)], col = iris.class)
```
CHALLANGE 2

```{r}
salary <- read.csv('http://bit.ly/SalaryDatasetClustering')
```


```{r}
head(salary)
```

```{r}
dim(salary)
```


```{r}
# null values
sum(is.na(salary))
```


```{r}
summary(salary)
```


```{r}
df <- select(salary,-c(Status,Notes,Benefits, EmployeeName))
```

```{r}
str(salary)
```


```{r}
# checking for unique values
unique(salary$Status)
unique(salary$Notes)
unique(salary$Benefits)
```


```{r}
for (x in salary$Status){
    if (x == "PT"){
        print(count(x))
    }  
    
}
```


```{r}
df$Agency <- as.integer(as.factor(df$Agency))
df$JobTitle <- as.integer(as.factor(df$JobTitle))
```


```{r}
head(df)
```


```{r}
summary(df)
```


```{r}
# scale the data

normalize <- function(x){
  return ((x-min(x)) / (max(x)-min(x)))
}

df$JobTitle <- normalize(df$JobTitle)
df$TotalPay <- normalize(df$TotalPay)
df$TotalPayBenefits <- normalize(df$TotalPayBenefits)
df$Year <- normalize(df$Year)
```


```{r}
sum(is.na(df))
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```

