---
title: "linear reg"
author: "Faith"
date: "9/8/2020"
output: html_document
---

```{r}
library(tidyverse) # data manipulation and viz
library(modelr) # provides easy pipeline modeling function
library(broom) # helps to tidy up model outputs
```
```{r}
# load the data
getwd()
setwd("C:/Users/FGakori/Documents/supervised and unsupervised")

advert <- read.csv('advert.csv', header = TRUE)
advert
```
preparing the data
use 60% - 40% split
```{r}
set.seed(123)
sample <- sample(c(TRUE, FALSE), nrow(advert), replace = T, prob = c(0.6,0.4))

train <- advert[sample, ]
test <- advert[!sample, ]
```


# simple linear regression
tv budget abd sales

## model building
```{r}
model1 <- lm(Sales ~ TV, data = train)
```

```{r}
summary(model1)
```

print tidt version of the model

```{r}
tidy(model1)
```
6.86 is the intercept abd 0.05 is the slope(TV)

Stat significance of the coefficients
```{r}
confint(model1)
```

# Assessing Model Auracy

RSE is an estimate of the standars deviation.average amount that the response will deviate from the true regression line
```{r}
sigma(model1)
```

R^2 represents the proportion of variance explained and so it always takes on a value btwn 0 and 1.
```{r}
rsquare(model1, data = train)

```

F-stat tests to see if at least one predictor variable has a non-zero coefficient.

# making predictions

```{r}
(test <- test %>%
   add_predictions(model1))
```

# imputing MSE
```{r}
#test MSE
test %>% 
  add_predictions(model1) %>%
  summarise(MSE = mean((Sales - pred)^2))
```
```{r}
# train MSE
train %>% 
  add_predictions(model1) %>%
  summarise(MSE = mean((Sales - pred)^2))
```
# MULTIPLE REGRESSION
## Model building

```{r}
model2 <- lm(Sales ~ TV + Radio + Newspaper, data = train)
```

```{r}
summary(model2)
```
changes in newspaper do not appear to have a relationship witj changes in sales as th pvalue is greater than 0.05

```{r}
tidy(model2)
```

# model accuracy

```{r}
list(model1 = broom::glance(model1), model2 = broom::glance(model2))
```


# height and weight data

```{r}
x <- c(151, 174, 138, 186, 128, 136, 179, 163, 152, 131)
y <- c(63, 81, 56, 91, 47, 57, 76, 72, 62, 48)
```


apply lm() function

```{r}
relation <- lm(y~x)
print(relation)
```

```{r}
summary(relation)
```

predict() function

```{r}
# Find weight of a person with height 170.

a <- data.frame(x = 170)
result <-  predict(relation,a)
print(result)
```

